### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Working on Chapter 4, I want to have a first correct version by tomorrow.

\TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}

@@

As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol.
All devices share a synchronized time, and know in advance the finite number of available RF channels.
In each time slot, devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (no sensing).
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station. If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part.
If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \emph{Ack} is not transmitted.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.28]{2-Chapters/4-Chapter/CrownCom_17.git/protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets. If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

@@

There are $N_c \geq 1$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
Furthermore, in channel $i \in \llbracket 1; N_c \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{N_c} S_i$).
We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $N_c$ (about $1000$ to $10000$, while $N_c$ is about $10$ to $50$).
As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq N_c$, as the former makes more sense for IoT networks\footnote{Here, $N_c$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.

We suppose that all devices follow the same emission pattern, being fixed in time, and we choose to model it as a simple Bernoulli process:
all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $N_c / (S + D)$ for all devices to communicate successfully (in average).}.
The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed (\ie, real time during two messages), and $1/p$ is proportional to the \emph{duty cycle}.
For instance for IoT objects, a ``smart'' refrigerator could have to send a daily message, and if a message can be sent every second, then $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.

@@

We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; N_c \rrbracket = \{1, \dots, N_c\}$).
% every time they communicate
% This reflects a very naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.

In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $i$, at that time slot.
The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$ :
\begin{equation}
    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{N_c} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/N_c, \text{by uniform choice}}
\end{equation}

All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $i$ for any dynamic device is $p / N_c$, and there are $D-1$ other dynamic devices.
As they are independent, the probability that no other dynamic device sent in $i$
% $\Pr(\text{no other dynamic device sent in}\;i)$,
is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device}\;k\;\text{did not sent in}\;i) = \prod_{k=1}^{D-1} \Pr(\text{device}\;k\;\text{did not sent in}\;i)$. And $\Pr(\text{device}\;k\;\text{sent in}\;i) = p \times 1 / N_c$, by uniform choice on channels and the Bernoulli emission hypothesis. So $q = \prod_{k=1}^{D-1} (1 - p/N_c) = (1-p/N_c)^{D-1}$. Thus we can conclude,
%
\begin{align}\label{eq:41:strategynaive}
    \Pr(\text{success}|\text{sent})
    & = \sum_{i=1}^{N_c} \underbrace{(1 - p / N_c)^{D-1}}_{\text{No other dynamic device}} \times \underbrace{(1-p)^{S_i}}_{\text{No static device}} \times\; \frac{1}{N_c} \nonumber \\
    & = \frac{1}{N_c} \left(1-\frac{p}{N_c}\right)^{D-1} \sum_{i=1}^{N_c} (1-p)^{S_i} .
\end{align}
This expression \eqref{eq:41:strategynaive} is constant (in time), and easy to compute numerically, but comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it.

@@

If the oracle draws once uniformly at random a configuration of dynamic devices, with $D_i$ devices affected to channel $i$ is fixed (in time, \ie, stationary),
then this probability is computed as before:
\begin{align}\label{eq:41:prob_col}
    \Pr(\text{success}|\text{sent})
    & = \sum_{i=1}^{N_c} \Pr(\text{success}|\text{sent in channel}\;i) \; \Pr(\text{sent in channel}\;i) \nonumber \\
    & = \sum_{i=1}^{N_c} \underbrace{(1 - p)^{D_i - 1}}_{\;\;D_i - 1 \;\text{others}\;\;} \times \underbrace{(1 - p)^{S_i}}_{\;\;\text{No static device}\;\;} \times \underbrace{ D_i / D }_{\;\;\text{Sent in channel}\; i\;\;}.
\end{align}

Consequently, the optimal allocation vector $(D_1,\dots,D_{N_c})$ is the solution of the following real-valued constraint optimization problem :
\begin{subequations}
\label{eq:41:prob}
\begin{align}
\underset{D_1,\dots,D_{N_c}}{\arg\max}\; & \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
\text{such that}\;\; & \sum_{i=1}^{N_c} D_i = D, \label{eq:41:eqCstr}\\
& D_i \geq 0 \qquad \forall i\in\llbracket 1;N_c\rrbracket . \label{eq:41:ineqCstr}
\end{align}
\end{subequations}

@@

\begin{equation}\label{eq:41:Dilambda}
D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{\dag} .
\end{equation}
\end{proposition}
\begin{proof}
\begin{itemize}
    \item
    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;N_c \rrbracket$. For such values for $D_i$, the objective function $f: (D_1, \dots, D_{N_c}) \mapsto \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{N_c}$ \cite{Luenberger68,Yaari77}.}.
        \item
The Lagrange multipliers method can be applied to the optimization problem \eqref{eq:41:optPb}, with a concave objective function $f$, linear equality constraints \eqref{eq:41:eqCstr} and linear inequality constraints \eqref{eq:41:ineqCstr}. The strong duality condition is satisfied in this case \cite{BoydVanderberghe04}, so finding the saddle points will be enough to find the maximizers.
    \hfill{}$\square$
    \end{itemize}
\end{proof}

Where $(a)^{\dag} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto xe^x$ on $\mathbb{R^+} = [0, +\infty)$ \cite{Corless96}.
Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint
\begin{equation}\label{eq:41:constraintLambda}
 \sum_{i=1}^{N_c} D_i^*(\lambda) = D.
\end{equation}

Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded to find the optimal number of devices for channel $i$ :
% Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < N_c$, and $\widehat{D_{N_c}} = D - \sum_{i=1}^{N_c - 1} \widehat{D_i}$.

@@

% A Multi-Armed Bandit problem is defined as follows \cite{Thompson33,Robbins52,LaiRobbins85}.
% There is a fixed number $N_c \geq 1$ of levers, or ``arms'', and a player has to choose one lever at each discrete time $t \geq 1, t \in \mathbb{N}$, denoted as $A(t) = k \in\{1,\dots,N_c\}$.
% Selecting arm $k$ at time $t$ yields a (random) \emph{reward}, $r_k(t) \in \mathbb{R}$, and the goal of the player is to maximize the sum of his rewards, $r_{1 \dots T} = \sum_{t = 1}^T r_{A(t)}(t)$.

@@

% Simulation parameters
We suppose a network with $S + D = 2000$ end-devices, and one IoT base station.
Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\emph{e.g.}, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
The RF band is divided in $N_c = 10$ channels.
Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{N_c}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases. The dynamic devices have access to all the channels, and use learning algorithms.
%to find the least loaded.
%
We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
We tried similar experiments with other values for $N_c$ and this repartition vector, and results were similar for non-homogeneous repartitions. Clearly, the problem is less interesting for homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Fig.\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.

>>>>

% \TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}

@@

As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
All devices share a synchronized time, and know in advance the finite number of available RF channels.
In each time slot, devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (no sensing).
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station. If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part.
If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \emph{Ack} is not transmitted.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.50]{2-Chapters/4-Chapter/CrownCom_17.git/protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets. If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

@@

There are $K \geq 1$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
Furthermore, in channel $i \in \llbracket 1; K \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{K} S_i$).
We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (about $1000$ to $10000$, while $K$ is about $10$ to $50$).
As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks\footnote{Here, $K$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.

We suppose that all devices follow the same emission pattern, being fixed in time, and we choose to model it as a simple Bernoulli process:
all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $K / (S + D)$ for all devices to communicate successfully (in average).}.
The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed (\ie, real time during two messages), and $1/p$ is proportional to the \emph{duty cycle}.
For instance for IoT objects, a ``smart'' refrigerator could have to send a daily message, and if a message can be sent every second, then $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.

@@

We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; K \rrbracket = \{1, \dots, K\}$).
% every time they communicate
% This reflects a very naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.

In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $i$, at that time slot.
The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$ :
\begin{equation}
    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/K, \text{by uniform choice}}
\end{equation}

All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $i$ for any dynamic device is $p / K$, and there are $D-1$ other dynamic devices.
As they are independent, the probability that no other dynamic device sent in $i$
% $\Pr(\text{no other dynamic device sent in}\;i)$,
is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device}\;k\;\text{did not sent in}\;i) = \prod_{k=1}^{D-1} \Pr(\text{device}\;k\;\text{did not sent in}\;i)$. And $\Pr(\text{device}\;k\;\text{sent in}\;i) = p \times 1 / K$, by uniform choice on channels and the Bernoulli emission hypothesis. So $q = \prod_{k=1}^{D-1} (1 - p/K) = (1-p/K)^{D-1}$. Thus we can conclude,
%
\begin{align}\label{eq:41:strategynaive}
    \Pr(\text{success}|\text{sent})
    & = \sum_{i=1}^{K} \underbrace{(1 - p / K)^{D-1}}_{\text{No other dynamic device}} \times \underbrace{(1-p)^{S_i}}_{\text{No static device}} \times\; \frac{1}{K} \nonumber \\
    & = \frac{1}{K} \left(1-\frac{p}{K}\right)^{D-1} \sum_{i=1}^{K} (1-p)^{S_i} .
\end{align}
This expression \eqref{eq:41:strategynaive} is constant (in time), and easy to compute numerically, but comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it.

@@

If the oracle draws once uniformly at random a configuration of dynamic devices, with $D_i$ devices affected to channel $i$ is fixed (in time, \ie, stationary),
then this probability is computed as before:
\begin{align}\label{eq:41:prob_col}
    \Pr(\text{success}|\text{sent})
    & = \sum_{i=1}^{K} \Pr(\text{success}|\text{sent in channel}\;i) \; \Pr(\text{sent in channel}\;i) \nonumber \\
    & = \sum_{i=1}^{K} \underbrace{(1 - p)^{D_i - 1}}_{\;\;D_i - 1 \;\text{others}\;\;} \times \underbrace{(1 - p)^{S_i}}_{\;\;\text{No static device}\;\;} \times \underbrace{ D_i / D }_{\;\;\text{Sent in channel}\; i\;\;}.
\end{align}

Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem :
\begin{subequations}
\label{eq:41:prob}
\begin{align}
\underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
\text{such that}\;\; & \sum_{i=1}^{K} D_i = D, \label{eq:41:eqCstr}\\
& D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
\end{align}
\end{subequations}

@@

\begin{equation}\label{eq:41:Dilambda}
D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{\dag} .
\end{equation}
\end{proposition}
\begin{proof}
\begin{itemize}
    \item
    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;K \rrbracket$. For such values for $D_i$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
        \item
The Lagrange multipliers method can be applied to the optimization problem \eqref{eq:41:optPb}, with a concave objective function $f$, linear equality constraints \eqref{eq:41:eqCstr} and linear inequality constraints \eqref{eq:41:ineqCstr}. The strong duality condition is satisfied in this case \cite{BoydVanderberghe04}, so finding the saddle points will be enough to find the maximizers.
    \hfill{}$\square$
    \end{itemize}
\end{proof}

Where $(a)^{\dag} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto xe^x$ on $\mathbb{R^+} = [0, +\infty)$ \cite{Corless96}.
Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint
\begin{equation}\label{eq:41:constraintLambda}
 \sum_{i=1}^{K} D_i^*(\lambda) = D.
\end{equation}

Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded to find the optimal number of devices for channel $i$ :
% Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < K$, and $\widehat{D_{K}} = D - \sum_{i=1}^{K - 1} \widehat{D_i}$.

@@

% A Multi-Armed Bandit problem is defined as follows \cite{Thompson33,Robbins52,LaiRobbins85}.
% There is a fixed number $K \geq 1$ of levers, or ``arms'', and a player has to choose one lever at each discrete time $t \geq 1, t \in \mathbb{N}$, denoted as $A(t) = k \in\{1,\dots,K\}$.
% Selecting arm $k$ at time $t$ yields a (random) \emph{reward}, $r_k(t) \in \mathbb{R}$, and the goal of the player is to maximize the sum of his rewards, $r_{1 \dots T} = \sum_{t = 1}^T r_{A(t)}(t)$.

@@

% Simulation parameters
We suppose a network with $S + D = 2000$ end-devices, and one IoT base station.
Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\emph{e.g.}, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
The RF band is divided in $K = 10$ channels.
Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{K}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases. The dynamic devices have access to all the channels, and use learning algorithms.
%to find the least loaded.
%
We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions. Clearly, the problem is less interesting for homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Fig.\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Working on Chapter 4, I want to have a first correct version by tomorrow.

\subsection{A missing proof}

This short subsection contains the missing details of the proof of Proposition \ref{prop:41:Lagrangian}.
%
First, we need to justify that the objective function is quasi-concave, in each of its coordinates.
%
Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($N_c, p$), the repartition of static devices ($S_1,\dots,S_{N_c}$) and the Lagrange multiplier $\lambda$.

@@

    \item
    The function $f: (D_1, \dots, D_{N_c}) \mapsto \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{N_c}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
    \hfill{}$\square$
\end{itemize}

@@

Let us now prove the expression for $D_i^*$ given in \eqref{eq:41:Dilambda}.
\begin{itemize}
	\item
	If $\boldsymbol{D} = (D_1,\dots,D_{N_c})$,
    the Lagrangian is denoted $\mathcal{L}(\boldsymbol{D}, \lambda) = f(\boldsymbol{D}) + \lambda(D - \sum_{i=1}^{N_c} D_i)$, and its derivative w.r.t. $D_i$ is
    $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) = (1-p)^{S_i + D_i - 1} + \log(1 - p) D_i (1-p)^{S_i + D_i - 1} - \lambda$.

>>>>

\subsection{Proof of Proposition \ref{prop:41:Lagrangian}}

This short subsection contains the missing details of the proof of Proposition \ref{prop:41:Lagrangian}.
%
First, we need to justify that the objective function is quasi-concave, in each of its coordinates.
%
Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($K, p$), the repartition of static devices ($S_1,\dots,S_{K}$) and the Lagrange multiplier $\lambda$.

@@

    \item
    The function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{K}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
    \hfill{}$\square$
\end{itemize}

@@

Let us now prove the expression for $D_i^*$ given in \eqref{eq:41:Dilambda}.
\begin{itemize}
	\item
	If $\boldsymbol{D} = (D_1,\dots,D_{K})$,
    the Lagrangian is denoted $\mathcal{L}(\boldsymbol{D}, \lambda) = f(\boldsymbol{D}) + \lambda(D - \sum_{i=1}^{K} D_i)$, and its derivative w.r.t. $D_i$ is
    $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) = (1-p)^{S_i + D_i - 1} + \log(1 - p) D_i (1-p)^{S_i + D_i - 1} - \lambda$.

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
Ok for tex root
Okay for chapters and parts
Working on Chapter 4, I want to have a first correct version by tomorrow.

%!TEX root = ../thesis.tex

% First chapter begins here
\chapter{Multi-Armed Bandits Models for Internet of Things Networks}
\label{chapter:4}
\minitoc
\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/4-Chapter/Images/}}

@@

This chapter has three sections, each is a raw include from different papers

- ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419

- ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825

- ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}

% ----------------------------------------------------------------------------
\section{A first model of learning devices in an IoT network}
\label{sec:4:firstModel}

@@

% ----------------------------------------------------------------------------
\section{Conclusion of Chapter 4}
\label{sec:4:conclusion}

@@

In this chapter, we saw...

Future works include...

% ----------------------------------------------------------------------------
\section{Appendix for Chapter 4}
\label{sec:4:appendix}

>>>>

%!TEX root = ../PhD_thesis__Lilian_Besson

\chapter{More Efficient Spectrum Usage of IoT Networks with Selfish MAB Learning}
\label{chapter:4}
\minitoc

In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithm to learn and improve their access to the network.
The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP \cite{USRPDocumentation} cards.

More precisely, on the one hand, we start by presenting in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, used by static devices.
Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their communication messages were successfully received.
We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time. If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
we show in simulations that in the case of non-uniform repartition of static devices in the channels, the dynamic devices can improve a lot the network efficiency.
%
On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real radio hardware.

On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending the first one, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
In this second model, we compare different heuristic, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.

\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/4-Chapter/Images/}}

@@

\paragraph{Publications}
This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

% ----------------------------------------------------------------------------
\section{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
\label{sec:4:firstModel}

@@

% ----------------------------------------------------------------------------
\section{Conclusion of Chapter~\ref{chapter:4}}
\label{sec:4:conclusion}

@@

\paragraph{Summary}
In this chapter, we saw...

\paragraph{Future works}
Future works include...

% ----------------------------------------------------------------------------
\section{Appendix for Chapter~\ref{chapter:4}}
\label{sec:4:appendix}
